### 2019-01-19 metrics2.tex
Metrics

\begin{enumerate}[(i)]
	\item Since 
	\begin{align*}
	\tilde{\beta_1} = a_1 y_1 + \dots + a_ny_n
	\end{align*}
	is a linear function of $y_1, \dots, y_n$ with some coefficients, thus, it is a linear estimator. Now let us show unbiasedness  
	\begin{align*}
	\tilde{\beta_1} = \frac{\sum_{i=1}^n (z_i - \bar{z})(\beta_0 + \beta_1 x_i + u_i)}{\sum_{i=1}^n(z_i - \bar{z})x_i} = \beta_1 + \frac{\sum_{i=1}^n(z_i - \bar{z})u_i}{\sum_{i=1}^n(z_i - \bar{z})x_i}\\
	\expect \tilde{\beta_1} = \beta_1 + \expect \left[\frac{\sum_{i=1}^n(z_i - \bar{z})u_i}{\sum_{i=1}^n(z_i - \bar{z})x_i}\right]
	\end{align*}
	\begin{align*}
	\expect[(z_i - \bar{z})u_i] \underset{\expect u_i = 0}{ = } \expect[z_iu_i] = \expect[\expect[z_iu_i|x_1, \dots, x_n]] \underset{\text{i.i.d}}{ = } \expect[z_i\expect[u_i|x_i]] \underset{\text{G.-M. assump.}}{ = } 0 
	\end{align*}
	Hence 
	\begin{align*}
	\expect \tilde{\beta_1} = \beta_1
	\end{align*}
	i.e. the estimator is unbiased.
	\item Assume $Var(u|x) = \sigma^2$ then
	\begin{align*}
	Var \tilde{\beta_1} &= Var\left[\frac{\sum_{i=1}^n (z_i - \bar{z})y_i}{\sum_{i=1}^n (z_i - \bar{z})x_i}\right] = \frac{1}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2} Var\left[\sum_{i=1}^n (z_i - \bar{z})y_i\right] \underset{\text{i.i.d}}{=}\\
	&\underset{\text{i.i.d}}{=}\frac{1}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2} \sum_{i=1}^n Var((z_i - \bar{z})y_i)\\
	&Var((z_i - \bar{z})y_i) = \expect[Var[((z_i - \bar{z})y_i)|x_1, \dots, x_n]] + Var(\expect[((z_i - \bar{z})y_i)|x_1, \dots, x_n]) = \\
	&=\expect[(z_i - \bar{z})^2 Var(y_i|x_1, \dots, x_n)] + Var((z_i - \bar{z})\expect[y_i|x_1, \dots, x_n]) \underset{\text{i.i.d}}{=}\\
	&\underset{\text{i.i.d}}{=} \expect[(z_i-\bar{z})^2 Var(y_i|x_i)] + Var((z_i - \bar{z})\expect[y_i|x_i])
	\end{align*}
	\begin{align*}
	Var(y_i|x_i) = Var(\beta_0 + \beta_1x_i + u_i|x_i) = Var(u_i|x_i) = \sigma^2\\
	\expect[y_i|x_i] = \expect[\beta_0 + \beta_1x_i + u_i|x_i] = \beta_0 + \beta_1x_i
	\end{align*}
	Hence
	\begin{align*}
	Var ((z_i - \bar{z})y_i) = (z_i - \bar{z})^2\sigma^2 + Var(const) = (z_i - \bar{z})^2\sigma^2
	\end{align*}
	Thus, 
	\begin{align*}
	Var \tilde{\beta_1} = \sigma^2 \frac{\sum_{i=1}^n (z_i -  \bar{z})^2}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2}
	\end{align*}
	\item 
	\begin{align*}
	\hat{\beta_1} = \frac{\sum_{i=1}^n(x_i - \bar{x})y_i}{\sum_{i=1}^n(x_i - \bar{x})^2}
	\end{align*}
	Similarly to (ii):
	\begin{align*}
	Var \hat{\beta_1} = \frac{ \sigma^2}{\sum_{i=1}^n(x_i - \bar{x})^2}
	\end{align*}
	By the Cauchy - Schwartz inequality:
	\begin{align*}
	\left(\sum_{i=1}^n (z_i - \bar{z})(x_i - \bar{x})\right)^2 = \left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2 \le \left(\sum_{i=1}^n (x_i - \bar{x})^2\right) \left(\sum_{i=1}^n (z_i - \bar{z})^2\right)
	\end{align*} 
	hence
	\begin{align*}
	Var \tilde{\beta_1} \ge \sigma^2 \frac{\summa (z_i - \bar{z})^2}{\left(\sum_{i=1}^n (x_i - \bar{x})^2\right) \left(\sum_{i=1}^n (z_i - \bar{z})^2\right)} = \frac{ \sigma^2}{\sum_{i=1}^n(x_i - \bar{x})^2} = Var \hat{\beta_1}
	\end{align*}
\end{enumerate}
\end{document}

>>>>

\begin{enumerate}[(i)]
	\item Since 
	\begin{align*}
	\tilde{\beta_1} = a_1 y_1 + \dots + a_ny_n
	\end{align*}
	is a linear function of $y_1, \dots, y_n$ with some coefficients, thus, it is a linear estimator. Now let us show unbiasedness  
	\begin{align*}
	\tilde{\beta_1} = \frac{\sum_{i=1}^n (z_i - \bar{z})(\beta_0 + \beta_1 x_i + u_i)}{\sum_{i=1}^n(z_i - \bar{z})x_i} = \beta_1 + \frac{\sum_{i=1}^n(z_i - \bar{z})u_i}{\sum_{i=1}^n(z_i - \bar{z})x_i}\\
	\expect \tilde{\beta_1} = \beta_1 + \expect \left[\frac{\sum_{i=1}^n(z_i - \bar{z})u_i}{\sum_{i=1}^n(z_i - \bar{z})x_i}\right]
	\end{align*}
	\begin{align*}
	\expect[(z_i - \bar{z})u_i] \underset{\expect u_i = 0}{ = } \expect[z_iu_i] = \expect[\expect[z_iu_i|x_1, \dots, x_n]] \underset{\text{i.i.d}}{ = } \expect[z_i\expect[u_i|x_i]] \underset{\text{G.-M. assump.}}{ = } 0 
	\end{align*}
	Hence 
	\begin{align*}
	\expect \tilde{\beta_1} = \beta_1
	\end{align*}
	i.e. the estimator is unbiased.
	\item Assume $Var(u|x) = \sigma^2$ then
	\begin{align*}
	Var \tilde{\beta_1} &= Var\left[\frac{\sum_{i=1}^n (z_i - \bar{z})y_i}{\sum_{i=1}^n (z_i - \bar{z})x_i}\right] = \frac{1}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2} Var\left[\sum_{i=1}^n (z_i - \bar{z})y_i\right] \underset{\text{i.i.d}}{=}\\
	&\underset{\text{i.i.d}}{=}\frac{1}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2} \sum_{i=1}^n Var((z_i - \bar{z})y_i)\\
	&Var((z_i - \bar{z})y_i) = \expect[Var[((z_i - \bar{z})y_i)|x_1, \dots, x_n]] + Var(\expect[((z_i - \bar{z})y_i)|x_1, \dots, x_n]) = \\
	&=\expect[(z_i - \bar{z})^2 Var(y_i|x_1, \dots, x_n)] + Var((z_i - \bar{z})\expect[y_i|x_1, \dots, x_n]) \underset{\text{i.i.d}}{=}\\
	&\underset{\text{i.i.d}}{=} \expect[(z_i-\bar{z})^2 Var(y_i|x_i)] + Var((z_i - \bar{z})\expect[y_i|x_i])
	\end{align*}
	\begin{align*}
	Var(y_i|x_i) = Var(\beta_0 + \beta_1x_i + u_i|x_i) = Var(u_i|x_i) = \sigma^2\\
	\expect[y_i|x_i] = \expect[\beta_0 + \beta_1x_i + u_i|x_i] = \beta_0 + \beta_1x_i
	\end{align*}
	Hence
	\begin{align*}
	Var ((z_i - \bar{z})y_i) = (z_i - \bar{z})^2\sigma^2 + Var(const) = (z_i - \bar{z})^2\sigma^2
	\end{align*}
	Thus, 
	\begin{align*}
	Var \tilde{\beta_1} = \sigma^2 \frac{\sum_{i=1}^n (z_i -  \bar{z})^2}{\left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2}
	\end{align*}
	\item 
	\begin{align*}
	\hat{\beta_1} = \frac{\sum_{i=1}^n(x_i - \bar{x})y_i}{\sum_{i=1}^n(x_i - \bar{x})^2}
	\end{align*}
	Similarly to (ii):
	\begin{align*}
	Var \hat{\beta_1} = \frac{ \sigma^2}{\sum_{i=1}^n(x_i - \bar{x})^2}
	\end{align*}
	By the Cauchy - Schwartz inequality:
	\begin{align*}
	\left(\sum_{i=1}^n (z_i - \bar{z})(x_i - \bar{x})\right)^2 = \left(\sum_{i=1}^n (z_i - \bar{z})x_i\right)^2 \le \left(\sum_{i=1}^n (x_i - \bar{x})^2\right) \left(\sum_{i=1}^n (z_i - \bar{z})^2\right)
	\end{align*} 
	hence
	\begin{align*}
	Var \tilde{\beta_1} \ge \sigma^2 \frac{\summa (z_i - \bar{z})^2}{\left(\sum_{i=1}^n (x_i - \bar{x})^2\right) \left(\sum_{i=1}^n (z_i - \bar{z})^2\right)} = \frac{ \sigma^2}{\sum_{i=1}^n(x_i - \bar{x})^2} = Var \hat{\beta_1}
	\end{align*}
\end{enumerate}
\section*{Problem 3}
Consider the standard linear multivariate regression model
\begin{align*}
y = \beta_0 + \beta_1x_1 + \beta_2x_2 + u
\end{align*}
under the Gauss-Markov assumptions.
\begin{enumerate}[(a)]
\item Derive the OLS estimate $\hat{\beta_2}$ in terms of sample variances and sample covariances of
random variables $x_1$, $x_2$ and $y$.
\item How would you get the OLS estimates $\hat{\beta_1}$ and $\hat{\beta_2}$ by computing a series of OLS estimates
from simple regressions only?
\item Suppose you have decided to get OLS estimates for $\gamma_0$, $\gamma_1$ and $\gamma_2$ from
\begin{align*}
x_2 = \gamma_0 + \gamma_1x_1 + \gamma_2 y + \nu.
\end{align*}
Is $\tilde{\beta_2} = \frac{1}{\hat{\gamma_2}}$ an unbiased estimate of $\beta_2$?
\end{enumerate}

\begin{enumerate}[(a)]
	\item Denote sample variances as $s^2_{ab}, a, b \in \left\{y, x_1, x_2\right\}$. To compute the OLS estimate we need to solve the following optimization problem:
	\begin{align*}
	\underset{b_0, b_1, b_2}{\min} \summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})^2\\
	\text{FOCs:} \begin{cases}
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i}) = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{1i} = 0\\
	\summa (y_i - b_0 - b_1x_{1i} - b_2x_{2i})x_{2i} = 0
	\end{cases}\\
	b_0 = \bar{y} - b_1\bar{x_1} - b_2 \bar{x_2}\\
	b_1\left(\summa x_{1i}^2\right) = \summa (y_i - \bar{y} + b_1\bar{x_1} + b_2\bar{x_2} - b_2x_{2i})x_{1i}
	\end{align*}
	\begin{align*}
	b_1 = \frac{\summa (y_i - \bar{y} - b_2(x_{2i} - \bar{x}))x_{1i}}{\summa (x_{1i} - \bar{x})^2}&=\frac{\summa (y_i - \bar{y})(x_{1i} - \bar{x_1}) - b_2\summa(x_{2i} - \bar{x_2})(x_{1i} - \bar{x_1})}{\summa (x_{1i} - \bar{x})^2} =\\
	&=\frac{s^2_{yx_1} - b_2s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Similarly for $b_2$:
	\begin{align}\label{eq2}
	b_2 = \frac{s^2_{yx_2} - b_1s^2_{x_1x_2}}{s^2_{x_2x_2}} &= \frac{s^2_{yx_2} - s^2_{yx_1} + b_2s^2_{x_1x_2}}{s^2_{x_2x_2}} \nonumber\\
	b_2 &= \frac{s^2_{yx_2} - s^2_{yx_1}}{s^2_{x_2x_2} - s^2_{x_1x_2}}
	\end{align}
	\item Performing simple linear regression of $y$ on $x_1$, $y$ on $x_2$, $x_1$ on $x_2$ and $x_2$ on $x_1$ denote OLS estimates for slope coefficients as $\alpha_1, \alpha_2, \alpha_3, \alpha_4$ respectively. It is known that
	\begin{align*}
	\alpha_1 = \frac{s^2_{yx_1}}{s^2_{x_1x_1}},\ \alpha_2 = \frac{s^2_{yx_2}}{s^2_{x_2x_2}},\ \alpha_3 = \frac{s^2_{x_1x_2}}{s^2_{x_2x_2}},\ \alpha_4 = \frac{s^2_{x_1x_2}}{s^2_{x_1x_1}}
	\end{align*}
	Divide numerator and denumerator of (\ref{eq2}) by $s^2_{x_1x_1}$ (assuming that it is not equal to 0):
	\begin{align*}
	b_2 = \frac{\alpha_2\frac{\alpha_4}{\alpha_3} - \alpha_1}{\frac{\alpha_4}{\alpha_3} - \alpha_4}
	\end{align*}
	Similarly for $b_1$:
	\begin{align*}
	b_1 = \frac{\alpha_1\frac{\alpha_3}{\alpha_4} - \alpha_2}{\frac{\alpha_3}{\alpha_4} - \alpha_3}
	\end{align*}
	\item From (\ref{eq2}):
	\begin{align*}
	\hat{\gamma_2} = \frac{s^2_{yx_2} - s^2_{x_1x_2}}{s^2_{yy} - s^2_{yx_1}} \to \frac{1}{\hat{\gamma_2}} = \frac{s^2_{yy} - s^2_{yx_1}}{s^2_{yx_2} - s^2_{x_1x_2}}
	\end{align*}
\end{enumerate}
\end{document}
